---
title: "Reproducible Research: Peer Assessment 1"
author: "Holly Parish"
date: "July 13, 2017"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


rdat <- read.csv("./activity.csv")
```

Remove the data that has na values and update format of date from factor to be a date format

```{r}

dat <- na.omit(rdat)
dat$date <- as.Date(dat$date)

```

##What is mean total number of steps taken per day?

```{r}
spd <- aggregate(steps ~ date, dat, sum)
hist(spd$steps, xlab = "NUmber of Steps per day", main = "Histrogram of total number of steps taken per day", col= "blue")
summary(spd)
```

##What is the average daily activity pattern?
Let's plot it on a graph
```{r}
avgspi <- aggregate(steps ~ interval, dat, mean)
plot(avgspi$interval,avgspi$steps, type='l', xlab = "5 min intervals", ylab = "average number of steps", main = "Average Number of Steps by Time Interval", col = "blue")
```

5-minute interval, on average across all the days in the dataset, with the maximum number of steps
```{r}
avgspi$interval[which.max(avgspi$steps)]
```

##Imputing missing values

```{r}
sum(is.na(rdat))


rdatc <- rdat
rdatc$date <- as.Date(rdatc$date)
rdatc$steps[is.na(rdatc$steps)] <- 0

spd2 <- aggregate(steps ~ date, rdatc, sum)

hist(spd2$steps, xlab = "Number of Steps per day", main = "Histrogram of total number of steps taken per day missing data assigned 0", col= "red")
summary(spd2)

```
By assigning the NA values to 0, the median and the mean dropped significantly compared to just removing them from the sample to calculate mean and median. For example - the mean went from 10766 to 9385 steps per day.

##Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lattice)
numdate <- as.POSIXlt(rdatc$date)$wday
newdat <- cbind(rdatc,numdate)
wEndOrDay <- ifelse(newdat$numdate == 6 | newdat$numdate ==0,"Weekend","Weekday")
wdat <- cbind(newdat,wEndOrDay)

avgswe <- aggregate(steps ~ date + wEndOrDay, wdat, mean)
xyplot(steps ~ date | wEndOrDay, avgswe, type = 'l', layout = c(1,2),xlab = "Date", ylabel= "Avg Num of Steps")
```
